# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Load the data using the following:

```{r}
data <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

First, here's a histogram showing the total number of steps each day:

```{r}

total <- tapply(data$steps,data$date,sum)

hist(total,main = "Number of Steps", xlab = "Total Steps per Day", 
     ylab = "Frequency")
```

The mean and median steps are given by:

```{r}
mean(total,na.rm=T)

median(total,na.rm=T)
```

## What is the average daily activity pattern?

```{r}
intervals <- as.numeric(labels(split(data$interval,data$interval)))
plot(intervals,tapply(data$steps,data$interval,mean, na.rm = T),
     type = "l",xlab = "Interval",ylab = "Mean Number of Steps")
```

The interval with the maximum average number of steps is interval 835, with a value of roughly 206.  Can be shown via the following:
```{r}
avg <- tapply(data$steps,data$interval,mean,na.rm=T)
avg[avg == max(avg)]
```

## Imputing missing values

Find total number of NAs:
```{r}
date.nas <- is.na(data$date)
intv.nas <- is.na(data$interval)
steps.nas <- is.na(data$steps)

NAs <- data.frame("Total Date NAs" = sum(date.nas), 
                  "Total Interval NAs" = sum(intv.nas),
                  "Total Step NAs" = sum(steps.nas))

library(xtable)
xt <- xtable(NAs)
print(xt,type="html")
```

Replace NAs with mean for that interval:
```{r}
data$steps[is.na(data$steps)] <- with(data, ave(steps, interval, 
   FUN = function(x) mean(x, na.rm = TRUE)))[is.na(data$steps)]
```

Make sure there are no NAs left in the steps column:
```{r}
sum(is.na(data$steps))
```


```{r}
total <- tapply(data$steps,data$date,sum)

hist(total,main = "Number of Steps", xlab = "Total Steps per Day", 
     ylab = "Frequency")
```

The mean and median steps are given by:

```{r}
mean(total)

median(total)
```

Imputing the NAs didn't seem to make a difference in the mean total value but did bring the median up 1 step.

## Are there differences in activity patterns between weekdays and weekends?

Create days column:
```{r}
data$days <- weekdays(as.Date(data$date))
```

Separate days into weekdays and weekends:

```{r}
data$day.type <- ifelse((data$days == 'Saturday' | data$days == 'Sunday'),
                        'weekend','weekday')
data$day.type <- factor(data$day.type)
```

Average steps across each interval for weekdays and weekends:
```{r}
avg <- aggregate(steps ~ interval + day.type, data = data, FUN= "mean" )
```

Plot the results:
```{r}
library(lattice)
xyplot(steps ~ interval|day.type, data = avg, type = 'l',layout = c(1,2))
```
